### 2025-07-10 actuarial.qmd
20250710 - quotes

> If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
> The equation might yield a probability of .90 that Professor X goes to the movie tonight.
> But if the family doctor announced that Professor X had just broken his leg, no sensible sociologist would stick with the equation.
> Why didn't the factor of 'broken leg' appear in the formula?
> Because broken legs are very rare, and in the sociologist's entire sample of 500 criterion cases plus 250 cross-validating cases, he did not come upon a single instance of it.
> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.'
> 
> --- Meehl [-@Meehl1957, p. 269–270]

>>>>

> *If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
> The equation might yield a probability of .90 that Professor X goes to the movie tonight.
> But if the family doctor announced that Professor X had just broken his leg, no sensible sociologist would stick with the equation.
> Why didn't the factor of 'broken leg' appear in the formula?
> Because broken legs are very rare, and in the sociologist's entire sample of 500 criterion cases plus 250 cross-validating cases, he did not come upon a single instance of it.
> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.'*
> 
> --- Meehl [-@Meehl1957, p. 269–270]

### 2025-07-10 base-rates.qmd
20250708 - track chunk execution times
20250710 - quotes

> A cab was involved in a hit-and-run accident at night.
> Two cab companies, the Green and the Blue, operate in the city.
> You are given the following data:
> 
> - 85% of the cabs in the city are Green and 15% are Blue.
> - A witness identified the cab as Blue.
> The court tested the reliability of the witness under the circumstances that existed on the night of the accident and concluded that the witness correctly identified each one of the two colors 80% of the time and failed 20% of the time.
> 
> What is the probability that the cab involved in the accident was Blue rather than Green?
> 
> --- Kahneman [-@Kahneman2011, p. 166]

>>>>

> *A cab was involved in a hit-and-run accident at night.
> Two cab companies, the Green and the Blue, operate in the city.
> You are given the following data:
> 
> - 85% of the cabs in the city are Green and 15% are Blue.
> - A witness identified the cab as Blue.
> The court tested the reliability of the witness under the circumstances that existed on the night of the accident and concluded that the witness correctly identified each one of the two colors 80% of the time and failed 20% of the time.
> 
> What is the probability that the cab involved in the accident was Blue rather than Green?*
> 
> --- Kahneman [-@Kahneman2011, p. 166]

### 2025-07-10 evaluating-prediction-accuracy.qmd
20250709 - videos
20250710 - video citations
20250710 - quotes

> "Nothing ruins fantasy more than reality." – [Renee Miller, Ph.D.](@fig-fantasy-vs-reality) [@Yahoo2024]

@@

> I consider it unnecessary to persuade you that most so-called "theories" in the soft areas of psychology (clinical, counseling, social, personality, community, and school psychology) are scientifically unimpressive and technologically worthless ... Perhaps the easiest way to convince yourself is by scanning the literature of soft psychology over the last 30 years and noticing what happens to theories.
> Most of them suffer the fate that General MacArthur ascribed to old generals—They never die, they just slowly fade away.
> In the developed sciences, theories tend either to become widely accepted and built into the larger edifice of well-tested human knowledge or else they suffer destruction in the face of recalcitrant facts and are abandoned, perhaps regretfully as a "nice try."
> But in fields like personology and social psychology, this seems not to happen.
> There is a period of enthusiasm about a new theory, a period of attempted application to several fact domains, a period of disillusionment as the negative data come in, a growing bafflement about inconsistent and unreplicable empirical results, multiple resort to ad hoc excuses, and then finally people just sort of lose interest in the thing and pursue other endeavors.
> 
> Meehl [-@Meehl1978, pp. 806–807]

>>>>

> "*Nothing ruins fantasy more than reality.*" – [Renee Miller, Ph.D.](#fig-fantasy-vs-reality) [@Yahoo2024]

@@

> *I consider it unnecessary to persuade you that most so-called "theories" in the soft areas of psychology (clinical, counseling, social, personality, community, and school psychology) are scientifically unimpressive and technologically worthless ... Perhaps the easiest way to convince yourself is by scanning the literature of soft psychology over the last 30 years and noticing what happens to theories.
> Most of them suffer the fate that General MacArthur ascribed to old generals—They never die, they just slowly fade away.
> In the developed sciences, theories tend either to become widely accepted and built into the larger edifice of well-tested human knowledge or else they suffer destruction in the face of recalcitrant facts and are abandoned, perhaps regretfully as a "nice try."
> But in fields like personology and social psychology, this seems not to happen.
> There is a period of enthusiasm about a new theory, a period of attempted application to several fact domains, a period of disillusionment as the negative data come in, a growing bafflement about inconsistent and unreplicable empirical results, multiple resort to ad hoc excuses, and then finally people just sort of lose interest in the thing and pursue other endeavors.*
> 
> --- Meehl [-@Meehl1978, pp. 806–807]

### 2025-07-10 intro.qmd
20250710 - quotes

> "*This material probably won't win you fantasy football championships.
> You could take what we learn and apply it to fantasy football and you might become 5 percent more likely to win.
> Or...
> Consider the broader relevance of this.
> You could learn data analysis and figure out ways to apply it to other systems.
> And you could be making a six-figure salary within the next five years.*" – [Benjamin Motz, Ph.D.](#fig-video-motz) [@NFL2014]

>>>>

> *This material probably won't win you fantasy football championships.
> You could take what we learn and apply it to fantasy football and you might become 5 percent more likely to win.
> Or...
> Consider the broader relevance of this.
> You could learn data analysis and figure out ways to apply it to other systems.
> And you could be making a six-figure salary within the next five years.*
> 
> --- [Benjamin Motz, Ph.D.](#fig-video-motz) [@NFL2014]

### 2025-07-10 player-evaluation.qmd
20250708 - track chunk execution times
20250710 - quotes

> "The best predictor of future behavior is past behavior." – Unknown

> "Past performance does not guarantee future results." – A common disclaimer about investments.

>>>>

> "*The best predictor of future behavior is past behavior.*" – Unknown

> "*Past performance does not guarantee future results.*" – A common disclaimer about investments.

### 2025-07-10 sports-cognitive-psychology.qmd
20250710 - quotes

> We work a menu of plays in that area over the course of the week.
> We rank them at the latter part of the week, and then we get into the stadium and we play it by ear.
> A lot of it has to do with the feel or the flow of the game—maybe what personnel group we think they're going to match our personnel group with.
> As we start to play and work the ball down the field on the drives that produce the touchdowns before the point after, we have a little inclination of what their personality might be at least in terms of matching our personnel.
> All of those things weigh into the decision.
> It legitimately is a feel thing.
> I think that's why you play the game.
> You can take analytics to baseball and things like that, but football is always going to be football.
> I got a lot of respect for analytics and numbers, but I'm not going to make judgments based on those numbers.
> The game is the game.
> It's an emotional one played by emotional and driven men.
> That's an element of the game you can't measure.
> Oftentimes, decisions such as that weigh heavily into the equation.
> 
> --- [Steelers Head Coach Mike Tomlin, 2015](https://steelersdepot.com/2015/09/tomlin-prefers-feel-over-analytics/) (archived at <https://perma.cc/7CHJ-BTWX>)

@@

> [Bruins Head Football Coach Kevin] Kelley believes that the "quant jocks" don't go far enough to validate the no-punting worldview and, more generally, the virtues of risk-taking.
> "The math guys, the astrophysicist guys, they just do the raw numbers and they don't figure emotion into it—and that's the biggest thing of all," he says.
> "The built-in emotion involved in football is unbelievable, and that's where the benefits really pay off."
> What he means is this: A defense that stops an opponent on third down is usually ecstatic.
> They've done their job.
> The punting unit comes on, and the offense takes over.
> When that defense instead gives up a fourth-down conversion, it has a hugely deflating effect.
> At Pulaski's games, you can see the shoulders of the opposing defensive players slump and their eyes look down when they fail to stop the Bruins on fourth down.
> 
> [@Moskowitz2011, p. 37]

>>>>

> "*We work a menu of plays in that area over the course of the week.
> We rank them at the latter part of the week, and then we get into the stadium and we play it by ear.
> A lot of it has to do with the feel or the flow of the game—maybe what personnel group we think they're going to match our personnel group with.
> As we start to play and work the ball down the field on the drives that produce the touchdowns before the point after, we have a little inclination of what their personality might be at least in terms of matching our personnel.
> All of those things weigh into the decision.
> It legitimately is a feel thing.
> I think that's why you play the game.
> You can take analytics to baseball and things like that, but football is always going to be football.
> I got a lot of respect for analytics and numbers, but I'm not going to make judgments based on those numbers.
> The game is the game.
> It's an emotional one played by emotional and driven men.
> That's an element of the game you can't measure.
> Oftentimes, decisions such as that weigh heavily into the equation.*"
> 
> --- [Steelers Head Coach Mike Tomlin, 2015](https://steelersdepot.com/2015/09/tomlin-prefers-feel-over-analytics/) (archived at <https://perma.cc/7CHJ-BTWX>)

@@

> [Bruins Head Football Coach Kevin] Kelley believes that the "*quant jocks*" don't go far enough to validate the no-punting worldview and, more generally, the virtues of risk-taking.
> "*The math guys, the astrophysicist guys, they just do the raw numbers and they don't figure emotion into it—and that's the biggest thing of all,*" he says.
> "*The built-in emotion involved in football is unbelievable, and that's where the benefits really pay off.*"
> What he means is this: A defense that stops an opponent on third down is usually ecstatic.
> They've done their job.
> The punting unit comes on, and the offense takes over.
> When that defense instead gives up a fourth-down conversion, it has a hugely deflating effect.
> At Pulaski's games, you can see the shoulders of the opposing defensive players slump and their eyes look down when they fail to stop the Bruins on fourth down.
> 
> --- [@Moskowitz2011, p. 37]
